Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2015-10-23 latex/article.tex
Final paper version

We first evaluate the curvature estimation on a full-resolution geometry
obtained by disabling the LoD criterion at the mesh generation step.
Figure~\ref{fig:full}-top shows results of curvature tensor
estimation (mean, Gaussian, first and
second principal directions) on an OctaFlower at level $l_0$ (ground truth).
Figure~\ref{fig:full}-bottom shows mean curvature estimation in real-time on a
dynamic object. For this case, we simply evaluate the implicit expression at
each vertex of a cell to decide if such cell is included or not into
$\Ball{R}{\vx}\cap \Shape$ instead of updating the \emph{mipmap texture} of
densities at each time step. For all illustrations, we use directly normal
vectors computed by algorithm discussed in Section~\ref{sec:curv-tens-estim}.

Then, we compare the approximation made by computing curvature from level $l \ge
l_0$ in Figure~\ref{fig:refine}. We can see that results using approximation
seems to fastly converge to the ground truth results.
Table~\ref{tab:full-res-stat} shows numerical results of $L_\infty$ error (worst
case) and $L_2$ error (mean), as well the number of \emph{mipmap texture}
fetches. We can also see that the number of texel fetch, required to compute the
curvature, reduces drastically when computing approximations. However, at higher levels, 
we get consequently much bigger errors. Those levels
are thus never used in practice, they are presented here to illustrate how the
refining process converges to the $l_0$ level. We see that levels close to $l_0$
lead very few error% than $l_0$, 
and should be considered for exploring our datasets
in real-time, changing parameters on-the-fly and visualizing dynamic datasets.

We also compare them with the hierarchical algorithm (as discussed in
Section~\ref{sec:hierarchical}). This method introduces a higher error when compared with
$l_0$. This is due to the precomputation of the subdivision that no longer
ensures to fetch data at the center of a mipmap cell. 
The GPU's interpolation reduces the bias, but it doesn't remove it.

@@

Those results were measured on a full resolution triangulation. 
However, since we compute the curvature of each vertices of the final
mesh, a highly triangulated object will be much slower to process.

@@

When dealing with large datasets, we can not expect real-time curvature tensor
estimation due to the huge amount of data to process. By computing a
dynamically refined approximate curvature tensor estimation joined with an
adaptive triangulation (as discussed in Section~\ref{sec:hierarchical}), we
manage to maintain a real-time framerate by giving control over the amount of
data to process at each frame. In Figure~\ref{fig:timings}, we compare timings
(in logarithmic scale) for a triangulation that is dynamically refining according
to its distance to the camera. We measured those timings using the regular and
the hierarchical algorithm.

@@

We can first note that the required time for an object to compute ground truth
curvature is usually as high as the time required to extract the geometry and
compute all of the above levels. Using approximations is thus mandatory.
Another advantage with approximation, is that it allows us to get a visualisation of our object as soon as we run
the application. This allows for real-time interactions with the object, required in order to change
the visualized value, curvature radius, \ldots.

It is also visible in Figure~\ref{fig:timings} that using a hierarchical decomposition greatly reduces the curvature computation
time, especially with big radii. 
%For the XYZ-Dragon with $R=20$, it needs almost
%4 seconds to compute the $l_0$ level, whereas using the hierarchical algorithm,
%it only needs 700ms. 
However, due to the precomputation and current hierarchical
structure, this algorithm is biased and creates an error (presented in
Table~\ref{tab:full-res-stat}) that needs to be considered.
%
% \begin{figure}
%   \vspace{-0.35cm}
%   \begin{center}
%     {\includegraphics[height=3.5cm]{figs/xyzrgb_dragon-510_R8_mean}}
%     {\includegraphics[height=3.0cm]{figs/snow_I08iso_233_r20_l0_m3}}
%   \end{center}
%   \caption{Illustration of mean curvature computation of two large datasets: XYZ-Dragon (with digital domain of $512^3$) and Snow microstructures ($233^3$).}
%   \label{fig:adaptive}
%   \vspace{-0.35cm}
% \end{figure}
%
% Figure~\ref{fig:adaptive} shows curvature computation and exploration in
% real-time on large datasets: XYZ-Dragon (with digital domain of $512^3$) and
% Snow microstructures ($233^3$).

@@

@@

An extension of this framework could be to implement an approximate hierarchical
algorithm. From there, we could use the dynamic refinement scheme to compute the
curvature with the hierarchical algorithm, thus massively reducing the number of
probes. We could also use texture streaming to compute on the fly on the CPU
the subdivision and stream it to the GPU, allowing dynamic radii change with the
hierarchical algorithm. Another improvement would be to implement the
hierarchical pattern presented in Figure\ref{fig:approx2}-$b$. This would prevent
bias from adding cells of density 1 but only partly convered by the
ball.

>>>>

We first evaluate the curvature estimation on a full-resolution geometry
obtained by disabling the LoD criterion at the mesh generation step.
Figure~\ref{fig:full}-top shows results of curvature tensor
estimation (mean, Gaussian, first and
second principal directions) on an OctaFlower at level $l_0$ (considered as our ground truth in the following).
Figure~\ref{fig:full}-bottom shows mean curvature estimation in real-time on a
dynamic object. For this case, we simply evaluate the implicit expression at
each vertex of a cell to decide if such cell is included or not into
$\Ball{R}{\vx}\cap \Shape$ instead of updating the \emph{mipmap texture} of
densities at each time step. For all illustrations, we use directly normal
vectors computed by algorithm discussed in Section~\ref{sec:curv-tens-estim}.

Then, we compare the approximation made by computing curvature from level $l \ge
l_0$ in Figure~\ref{fig:refine}. We can see that results using approximation
seems to quickly converge to the ground truth results.
Table~\ref{tab:full-res-stat} shows numerical results of $L_\infty$ error (maximal absolute difference for all vertices )
and $L_2$ error (mean absolute error of all vertices), as well the number of \emph{mipmap texture}
fetches. We can also see that the number of texel fetch, required to compute the
curvature, reduces drastically when computing approximations. However, at higher levels, 
approximation errors become more important. Those levels
are thus never used in practice, they are presented here to illustrate how the
refining process converges to the $l_0$ level. 
%We see that levels close to $l_0$
%lead very few error% than $l_0$, 
%and should be considered for exploring our datasets
%in real-time, changing parameters on-the-fly and visualizing dynamic datasets.

We also compare them with the hierarchical algorithm (as discussed in
Section~\ref{sec:hierarchical}). This method introduces a higher error when compared with
$l_0$. This is due to the precomputation of the subdivision that no longer
ensures to fetch data at the center of a mipmap cell. 
The GPU interpolation reduces the bias, but it doesn't remove it.

@@

%Those results were measured on a full resolution triangulation. 
%However, since we compute the curvature of each vertices of the final
%mesh, a highly triangulated object will be much slower to process.

@@

When dealing with large datasets, we can not expect real-time curvature tensor
estimation if we consider the full resolution geometry, due the the huge amount of data process. By computing a
dynamically refined approximate curvature tensor estimation joined with an
adaptive triangulation (as discussed in Section~\ref{sec:hierarchical}), we
manage to maintain a real-time framerate by giving control over the amount of
data to process at each frame. In Figure~\ref{fig:timings}, we compare timings
(in logarithmic scale) for a triangulation that is dynamically refined according
to its distance to the camera. We measured those timings using the multiresolution regular and
the hierarchical algorithm.

@@

We can first note that the required time for an object to compute the ground truth 
curvature is usually as high as the time required to extract the geometry and to
compute all of the above levels. Using approximations is thus mandatory.
Another advantage with approximations, is that it allows us to get a visualisation of our object as soon as we run
the application. This allows for real-time interactions with the object, required in order to change
the visualized value, curvature radius, \ldots

It is also visible in Figure~\ref{fig:timings} that a hierarchical decomposition greatly reduces the curvature computation
time, especially with big radii. 
%For the XYZ-Dragon with $R=20$, it needs almost
%4 seconds to compute the $l_0$ level, whereas using the hierarchical algorithm,
%it only needs 700ms. 
However, due to the precomputation and the current hierarchical
structure, this algorithm is biased and creates an error (presented in
Table~\ref{tab:full-res-stat}) that needs to be considered.

@@

 \begin{figure}[h!]
   \vspace{-0.35cm}
   \begin{center}
   	 {\includegraphics[width=0.45\textwidth]{figs/snow_r10}}
     {\includegraphics[width=0.5\textwidth]{figs/snow2_r10}}
     {\includegraphics[width=0.6\textwidth]{figs/xyzrgb_dragon-510_R8_mean}}
   \end{center}
   \caption{\emph{First row}: Mean curvature estimation on Snow Microstructures with a digital domain of $233^3$ and a radius of 10. 
   (left) Full snow object (right) exploration within the object
  \emph{Second row:} Mean curvature estimation on XYZ Dragon with a digital domain of $510^3$ and a radius of 8.}
   \label{fig:adaptive}
   \vspace{-0.35cm}
 \end{figure}

 Figure~\ref{fig:adaptive} shows curvature computation and exploration in
 real-time on large datasets: XYZ-Dragon (with digital domain of $512^3$) and
 Snow microstructures ($233^3$).

@@

%An extension of this framework could be to implement an approximate hierarchical
%algorithm. From there, we could use the dynamic refinement scheme to compute the
%curvature with the hierarchical algorithm, thus massively reducing the number of
%probes. We could also use texture streaming to compute on the fly on the CPU
%the subdivision and stream it to the GPU, allowing dynamic radii change with the
%hierarchical algorithm. Another improvement would be to implement the
%hierarchical pattern presented in Figure\ref{fig:approx2}-$b$. This would prevent
%bias from adding cells of density 1 but only partly convered by the
%ball.
